Incentivizing Exploration by Heterogeneous Users

### 2018-06-05 abstract.tex
edits to abstract and intro

We consider the problem of incentivizing exploration with heterogeneous agents.
In this problem, \ARMNUM bandit arms provide vector-valued outcomes equal to an unknown
arm-specific attribute vector,
perturbed by independent noise.
Agents arrive sequentially and choose arms to pull based on their own
private and heterogeneous linear utility functions over attributes
and the estimates of the arms' attribute vectors derived from
observations of other agents' past pulls.
Agents are myopic and selfish and thus would choose the arm with
maximum estimated utility.
A principal, knowing only the distribution from which agents'
preferences are drawn, but not the specific draws,
can offer incentive payments for pulling specific arms
to encourage agents to explore underplayed arms.
The principal seeks to minimize the total expected cumulative regret
incurred by agents relative to their best arms,
while also making a small expected cumulative payment.

We propose an algorithm that incentivizes infrequently played arms 
whose probability of being played in the next round would be small
without incentives.
Under the assumption that each arm is preferred by at
least a fraction $\MinProb > 0$ of agents,
we show that this algorithm achieves expected
cumulative regret of $O (\ARMNUM \e^{2/\MinProb} + \ARMNUM \log^3(T))$,
using expected cumulative payments of $O(\ARMNUM^2 \e^{2/\MinProb})$.
If \MinProb is known or the distribution over agent
preferences is discrete,
the exponential term $\e^{2/\MinProb}$ can be replaced with suitable
polynomials in \ARMNUM and $1/\MinProb$.
For discrete preferences, the regret's dependence on $T$ can be
eliminated entirely, giving constant (depending only polynomially on
\ARMNUM and $1/\MinProb$) expected regret and payments.
This constant regret stands in contrast to the $\Theta(\log(T))$ dependence of
regret in standard multi-armed bandit problems.
It arises because even unobserved heterogeneity in agent preferences
causes exploitation of arms to also explore arms fully;
succinctly, heterogeneity provides free exploration.

>>>>

We consider the problem of incentivizing exploration with heterogeneous agents.
In this problem, \ARMNUM bandit arms provide vector-valued outcomes equal to an unknown
arm-specific attribute vector,
perturbed by independent noise.
Agents arrive sequentially and choose arms to pull based on their own
private and heterogeneous linear utility functions over attributes
and the estimates of the arms' attribute vectors derived from
observations of other agents' past pulls.
Agents are myopic and selfish and thus would choose the arm with
maximum estimated utility.
A principal, knowing only the distribution from which agents'
preferences are drawn, but not the specific draws,
can offer \pfedit{arm-specific} incentive payments 
to encourage agents to explore underplayed arms.
The principal seeks to minimize the total expected cumulative regret
incurred by agents relative to their best arms,
while also making a small expected cumulative payment.

We propose an algorithm that incentivizes \pfedit{infrequently played arms}
whose probability of being played in the next round would be small
without incentives.
Under the assumption that each arm is preferred by at
least a fraction $\MinProb > 0$ of agents,
we show that this algorithm achieves expected
cumulative regret of $O (\ARMNUM \e^{2/\MinProb} + \ARMNUM \log^3(T))$,
using expected cumulative payments of $O(\ARMNUM^2 \e^{2/\MinProb})$.
If \MinProb is known or the distribution over agent
preferences is discrete,
the exponential term $\e^{2/\MinProb}$ can be replaced with suitable
polynomials in \ARMNUM and $1/\MinProb$.
For discrete preferences, the \pfedit{regret's} dependence on $T$ can be
eliminated entirely, giving constant (depending only polynomially on
\ARMNUM and $1/\MinProb$) expected regret and payments\pfdelete{where \ARMNUM is the number of arms}.
This constant regret stands in contrast to the $\Theta(\log(T))$ dependence of
regret in standard multi-armed bandit problems.
It arises because even unobserved heterogeneity in agent preferences
\pfedit{causes} exploitation of arms to also explore arms fully;
succinctly, heterogeneity provides free exploration.

### 2018-06-05 introduction.tex
edits to abstract and intro

Many websites and apps facilitate joint discovery,
sharing, and recommendation of content.
This includes news-, photo-, and video-sharing sites;
sites that host user-written reviews of products, restaurants, hotels, and travel experiences;
and citizen science projects
such as eBird \citep{sullivan2009ebird,xue-ebird} and Galaxy Zoo \citep{lintott-galaxy-zoo}.
Users of these sites improve their own experiences 
by learning from the experiences of others 
\citep{schmit2017human}.

Viewed more abstractly, users jointly explore a space of
options (products, news stories, photos, birdwatching sites,
%patches of sky to train their telescopes on, 
$\ldots$),
with the implicit goal of identifying the ``best'' ones.
Such scenarios are often modeled in a bandit
learning framework.
However, unlike standard bandit settings, the utilities of the
decision makers (the users) are not aligned with overall utility.
Societally (i.e., in a suitable aggregate over users),
substantial exploration of options could provide higher future rewards.
However, individual users interact with the site a limited number
of times, and therefore have little incentive to explore.
Recent work has focused on the setting where users interact only once, 
and so have no intrinsic incentive to explore.

Two recent lines of work have shown that
effecting a societally near-optimal outcome in this setting requires
explicitly incentivizing exploration:
\citet{kremer2014implementing}
and
\citet{mansour2015bayesian,mansour2016bayesian,mansour2018competition}
(see \cite{slivkins:asymmetry} for an overview)
assume that the site (also called the \emph{principal}) has an
informational advantage in being the only one to observe the results
of past arm pulls
(as in driving route recommendations).
The principal can use her%
\footnote{We use male pronouns to refer to users and female pronouns
  to refer to the principal.}
advantage to induce exploration by recommending apparently sub-optimal arms,
as long as agents cannot do better on their own.
\citet{frazier2014incentivizing} and 
\citet{han2015incentivizing} instead assume that the results of all
past arm pulls are publicly observable
(as on a review-sharing site).
They suppose the principal can incentivize exploration by offering payments to users as
a reward for pulling particular arms.
We follow this second model.
% We follow the model of \citet{frazier2014incentivizing} and
% \citet{han2015incentivizing} 
% and consider a multi-armed bandit model in which the principal can
% offer payments to the users for pulling particular arms.

Past work has assumed that users are homogeneous, i.e., the expected
reward derived from an arm is the same for all users.%
\footnote{\citet{han2015incentivizing} assume that users are
heterogeneous in their tradeoff between utility derived from arm pulls
and utility derived from the principal's payment.}
In reality, users have different preferences, e.g., gastronomic,
political, aesthetic, practical.
Indeed, the websites and mobile apps most widely used for joint discovery,
sharing, and recommendation of content tend to concern products and items 
with heterogeneity in preferences (movies, restaurants,
videos, travel experiences), and not items which have a universally
agreed-on best order.
This is perhaps because regimes with heterogeneous preferences are the
ones where discovery of the best items is the most difficult for people, 
and thus where online
platforms provide the most value.  Thus, we see an appropriate
accounting for heterogeneity as critical to an understanding of incentivizing
exploration in online communities.

Heterogeneity presents both a challenge and an opportunity.
On the one hand, unobserved heterogeneity hides critical
information about an agent's preferences from the principal.
On the other hand, heterogeneity also presents her an opportunity,
through the possibility of ``free exploration.''
Even when left unincentivized, agents will
play a variety of arms, revealing information about the arms' attributes.
This stands in sharp contrast to the case of homogeneous preferences,
where unincentivized agents will herd onto a single apparently best arm;
thus, effecting essentially any exploration at all requires incentives.

In order to take advantage of unobserved heterogeneity,
the principal has to give up some control, allowing the agent to
reveal his preferences through action, rather than obscuring them with
incentives.
However, she cannot give up control completely,
and must use incentives to force agents to explore against their
preferences, for the greater good.

With these challenges and opportunities in mind, our goal is to understand the
impact of user heterogeneity on the principal's ability to achieve
high social utility with low incentive payments,
and on the best approaches for doing so.
We wish to understand whether incentivizing exploration with
heterogeneous preferences  is ``harder'' or ``easier'' than with
homogeneous ones,
and to understand how exploration strategies that do well in the
heterogeneous preference setting differ from those in the homogeneous one.

Toward that end, we model our setting as follows.
(We describe our model at a high level here,
with formal definitions given in Section~\ref{sec:prob}.)
The \ARMNUM arms and users (or \emph{agents})
are characterized by payoff-relevant
\emph{attribute} (or \emph{feature}) vectors.
Arms' attributes are a priori unknown,
and agents' attributes are drawn from a known distribution.
An agent's reward from pulling an arm is the inner product of his
vector with the arm's vector (plus noise).
When an arm is pulled, a noisy version of its attribute vector is
observed by everyone.
This models full-text product reviews on websites like Amazon,
or ratings of ``service'', ``value'',
and other restaurant attributes on websites like Tripadvisor.  
Agents are myopic and will pull the arm whose expected attribute
vector (based on past noisy observations) maximizes their reward.
\dkedit{This assumption requires agents to have access to aggregate
information about the feature vector estimates of all (relevant) arms. 
Such access is facilitated by the platforms' ability to aggregate and
effectively display past feedback, e.g., by displaying simple averages
of ratings and using automatic summarization of full-text reviews 
\citep{wang2010product,liu2012movie,abulaish2009feature}.
While platforms could in principle manipulate the display of such
data, most have an incentive to be seen as truthful.}
%\pfedit{Agents' ability to effectively base decisions on all past observations models platforms' abilities to aggregate and effectively display past feedback, through simple averages of ratings and automatic summarization of full-text reviews \citep{wang2010product,liu2012movie,abulaish2009feature}.  While platforms can manipulate the display of user data, most have an incentive to be seen as truthful.}
The principal can incentivize agents to pull particular arms by
offering rewards for the specific arm.
The principal's goal is to keep the cumulative regret across all
agents small, while incurring only small total payments.
Our main theorem can be stated informally as follows:

@@

\begin{theorem} \label{thm:main-intro}
Assume that for each arm, at least a constant fraction \MinProb
of the population likes this arm best,
and let \TieDensity be a measure of the ``density of near-ties''
between agents' arm preferences
(in a sense made precise in Section~\ref{subsec:discrete}).
There is a policy that achieves expected 
cumulative regret $O (\ARMNUM \e^{2/\MinProb} + \dkedit{\TieDensity} \ARMNUM \log^3(T))$,
using expected cumulative payments of $O(\ARMNUM^2 \e^{2/\MinProb})$.
In particular, when agents who are close to tied between two arms have measure $0$,
both the expected regret and expected payment are bounded by constants
(with respect to $T$). 
\end{theorem}

\dkcomment{Added the \TieDensity term in the statement.}

The policy achieving the result of Theorem~\ref{thm:main-intro} is
quite simple: it mostly lets agents exploit arms, but incentivizes
them to explore when arms appear unlikely to be pulled without incentives.
It is presented in detail in Section~\ref{sec:ub}.

>>>>

\pfedit{
Many websites and apps facilitate joint discovery,
sharing, and recommendation of content.
This includes news-, photo-, and video-sharing sites;
sites that host user-written reviews of products, restaurants, hotels, and travel experiences;
and citizen science projects
such as eBird \citep{sullivan2009ebird,xue-ebird} and Galaxy Zoo \citep{lintott-galaxy-zoo}.
Users of these sites improve their own experiences 
by learning from the experiences of others 
\citep{schmit2017human}.
}

Viewed more abstractly, users jointly explore a space of
options (products, news stories, photos, birdwatching sites,
%patches of sky to train their telescopes on, 
$\ldots$),
with the implicit goal of identifying the ``best'' ones.
Such scenarios are often modeled in a bandit
learning framework.
However, unlike standard bandit settings, the utilities of the
decision makers (the users) are not aligned with overall utility.
Societally (i.e., in a suitable aggregate over users),
\pfedit{substantial exploration of options could provide higher future rewards.}
However, individual users interact with the site a limited number
of times, and therefore have little incentive to explore.
\pfedit{Recent work has focused on the setting where users interact only once, 
and so have no intrinsic incentive to explore.}

Two recent lines of work have shown that
effecting a societally \pfedit{near-optimal} outcome in this setting \pfedit{requires
explicitly inducing exploration}:
\citet{kremer2014implementing}
and
\citet{mansour2015bayesian,mansour2016bayesian,mansour2018competition}
(see \cite{slivkins:asymmetry} for an overview)
assume that the site (also called the \emph{principal}) has an
informational advantage in being the only one to observe the results
of past arm pulls
(as in driving route recommendations).
The principal can use her%
\footnote{We use male pronouns to refer to users and female pronouns
  to refer to the principal.}
\pfedit{advantage to induce exploration by recommending apparently sub-optimal arms,
as long as agents cannot do better on their own.}
\citet{frazier2014incentivizing} and 
\citet{han2015incentivizing} instead assume that the results of all
past arm pulls are publicly observable
\pfedit{(as on a review-sharing site)}.
They suppose the principal can \pfedit{incentivize exploration} by offering \pfedit{arm-specific reward payments}.
\pfedit{We follow this second model.}

Past work has assumed \pfedit{users have homogeneous preferences over arms, i.e., the expected
reward derived from an arm is the same for all users}.
(\citet{han2015incentivizing} \pfedit{models users that} are
heterogeneous in their tradeoff between utility derived from arm pulls
and utility derived from the principal's payment,
\pfedit{but assumes preferences across arms are homogenous.})
In reality, users have different preferences, e.g., gastronomic,
political, aesthetic, practical.
Indeed, the websites and mobile apps most widely used for joint discovery,
sharing, and recommendation of content tend to concern products and items 
with heterogeneity in preferences (movies, restaurants,
videos, travel experiences), and not items \pfedit{with} a universally
agreed-on best order.
This is perhaps because regimes with heterogeneous preferences are the
ones where \pfedit{people have the most difficulty discovering} the best items, 
and thus where online
platforms provide the most value.  
Thus, we see an appropriate
accounting for heterogeneity as critical to incentivizing
exploration in online communities.

\pfedit{In this work, we present the first algorithm and analysis (of which we
are aware) for incentivizing exploration when users have heterogeneous 
preferences over arms.}
\pfedit{Our analysis provides insight into the
impact of user heterogeneity on the principal's ability to achieve
high social utility with low incentive payments,
and on the best approaches for doing so.}
Heterogeneity presents both a challenge and an opportunity.
On the one hand, unobserved heterogeneity hides critical
information about an agent's preferences from the principal,
\pfedit{making her unaware of agents' preferred arms.
Thus, even with an unlimited incentive budget, incentivizing 
pulls according to a standard multi-armed bandit algorithm does not provide
low regret.  Instead, she must use incentives sparingly and 
allow agents to reveal their preferences through action.}
On the other hand, heterogeneity also \pfedit{offers}
the possibility of ``free exploration.''
Even \pfdelete{when left} unincentivized, agents will
play a variety of arms, revealing information about their attributes.
This stands in sharp contrast to the case of homogeneous preferences,
where unincentivized agents will herd onto a single apparently best arm,
\pfedit{and where} effecting essentially any exploration at all requires incentives.

\pfdelete{
With these challenges and opportunities in mind, 
our goal is to understand the
impact of user heterogeneity on the principal's ability to achieve
high social utility with low incentive payments,
and on the best approaches for doing so.
We wish to understand whether incentivizing exploration with
heterogeneous preferences  is ``harder'' or ``easier'' than with
homogeneous ones,
and to understand how exploration strategies that do well in the
heterogeneous preference setting differ from those in the homogeneous one.
Toward that end, we model our setting as follows.}
We describe our model at a high level here,
with formal definitions given in Section~\ref{sec:prob}.
The \ARMNUM arms and users (or \emph{agents})
are characterized by payoff-relevant
\emph{attribute} (or \emph{feature}) vectors.
Arms' attributes are a priori unknown,
and agents' attributes are drawn from a known distribution.
An agent's reward from pulling an arm is the inner product of his
vector with the arm's vector (plus noise).
When an arm is pulled, a noisy version of its attribute vector is
observed by everyone.
This models full-text product reviews on websites like Amazon,
or ratings of ``service'', ``value'',
and other restaurant attributes on websites like Tripadvisor.  
Agents are myopic and will pull the arm whose expected attribute
vector (based on past noisy observations) maximizes their reward.
This assumption requires agents to have access to aggregate
information about the feature vector estimates of all (relevant) arms. 
Such access is facilitated by the platforms' ability to aggregate and
effectively display past feedback, e.g., by displaying simple averages
of ratings and using automatic summarization of full-text reviews 
\citep{wang2010product,liu2012movie,abulaish2009feature}.
While platforms could in principle manipulate the display of such
data, most have an incentive to be seen as truthful.
%\pfedit{Agents' ability to effectively base decisions on all past observations models platforms' abilities to aggregate and effectively display past feedback, through simple averages of ratings and automatic summarization of full-text reviews \citep{wang2010product,liu2012movie,abulaish2009feature}.  While platforms can manipulate the display of user data, most have an incentive to be seen as truthful.}
The principal can incentivize agents to pull particular arms by
offering \pfedit{arm-specific payments}.
The principal's goal is to keep the cumulative regret across all
agents small, while incurring only small total payments.

Our main theorem can be stated informally as follows:

@@

\begin{theorem} \label{thm:main-intro}
Assume that for each arm, at least a constant fraction \MinProb
of the population likes this arm best,
and let \TieDensity be a measure of the ``density of near-ties''
between agents' arm preferences
(in a sense made precise in Section~\ref{subsec:discrete}).
There is a policy that achieves expected 
cumulative regret $O (\ARMNUM \e^{2/\MinProb} + \TieDensity \ARMNUM \log^3(T))$,
using expected cumulative payments of $O(\ARMNUM^2 \e^{2/\MinProb})$.
In particular, when agents who are close to tied between two arms have measure $0$,
both the expected regret and expected payment are bounded by constants
(with respect to $T$). 
\end{theorem}

The policy achieving the result of Theorem~\ref{thm:main-intro} is
simple: it mostly lets agents exploit arms, but incentivizes
them to explore when arms appear unlikely to be pulled without incentives.
It is presented in detail in Section~\ref{sec:ub}.

### 2018-06-05 problem_setting.tex
edits to abstract
edits to abstract and intro

We consider a multi-armed bandit setting with \ARMNUM arms.
Arm payoffs are determined by $d$ \emph{attributes} or \emph{features};
hence, arms are identified with vectors $\ArmV{i} \in \R^d$.
The \ArmV{i} are (adversarially) fixed, and unknown to the agents
or the principal.
When arm $i$ is pulled, its current utility-relevant features are
determined as $\ArmV{i} + \NoiseV$, where \NoiseV is a mean-zero independent Gaussian%
\footnote{For simplicity of notation, we assume that the variance
  $\sigma^2$ is uniform across time steps.
In fact, the analysis extends straightforwardly to any
mean-zero sub-Gaussian($\sigma^2$) noise that is independent but not necessarily identically distributed across time.}
noise vector $\NoiseV \sim \Normal{\vc{0}}{\sigma^2 I_{d}}$.
Here, $I_d$ denotes the $d \times d$ identity matrix.

At each time $t$, a new user (or \emph{agent}) arrives,
whose feature vector (which we also call his \emph{type})
$\AgentV{t} \in \R^d$ is drawn from a known distribution \AgentDist.
Depending on context, we will identify agents with their arrival time
$t$ or their type \AgentV{t}.
When agent $t$ pulls arm $\Pull{t} := i$,
he and all future agents observe a vector
$\ObsV{t} = \ArmV{i} + \NoiseV[t]$ for arm $i$,
and his reward is $\AgentV{t} \cdot \ObsV{t}$,
i.e., agents have linear preferences.
\pfedit{(We are also free to create new attributes that are nonlinear transformations of existing attributes.)}
\dkcomment{I have no idea what this sentence refers to.}

>>>>

We consider a multi-armed bandit setting with \ARMNUM arms.
Arm payoffs are determined by $d$ \emph{attributes} or \emph{features};
hence, arms are identified with vectors $\ArmV{i} \in \R^d$.
The \ArmV{i} are (adversarially) fixed, and unknown to the agents
and the principal.
When arm $i$ is pulled, its current utility-relevant features are
determined as $\ArmV{i} + \NoiseV$, where \NoiseV is a mean-zero independent Gaussian%
\footnote{For simplicity of notation, we assume that the variance
  $\sigma^2$ is uniform across time steps.
In fact, the analysis extends straightforwardly to any
mean-zero sub-Gaussian($\sigma^2$) noise that is independent but not necessarily identically distributed across time.}
noise vector $\NoiseV \sim \Normal{\vc{0}}{\sigma^2 I_{d}}$.
Here, $I_d$ denotes the $d \times d$ identity matrix.

At each time $t$, a new user (or \emph{agent}) arrives,
whose feature vector (which we also call his \emph{type})
$\AgentV{t} \in \R^d$ is drawn from a known distribution \AgentDist.
Depending on context, we will identify agents with their arrival time
$t$ or their type \AgentV{t}.
When agent $t$ pulls arm $\Pull{t} := i$,
he and all future agents observe a vector
$\ObsV{t} = \ArmV{i} + \NoiseV[t]$ for arm $i$,
and his reward is $\AgentV{t} \cdot \ObsV{t}$,
i.e., agents have linear preferences.
\pfedit{(Although linear, this model permits additional flexibility through the addition of derived attributes that are nonlinear transformation of some original attribute vector: define $\Agent{t}{i} = f_i(\AgentV{t})$ for some known nonlinear functions $f_i$ and $i=d+1,\ldots,d'$, and then increase $d$ to $d'$.)}
\pfcomment{I added some detail.  Is it clear now David?  I'm just talking about a standard trick in linear models, where you add nonlinear transformations of your feature vector and have a linear model in higher dimensions.}
